Distances

### 2020-04-08 03_distances.Rmd
Find nearest node and shortest paths on network

import numpy as np
import pandas as pd
```

@@

We use the `osmnx` library to get a highway network graph for New York City from 
OpenStreetMap. We

```{python graph}
B = ox.graph_from_place('New York, New York, USA', network_type='drive')
B_proj = ox.project_graph(B)
```

@@

# pad the county and tract strings, and get the proper data types
blockgroups['COUNTYFP'] = blockgroups['COUNTYFP'].apply('{0:0>3}'.format)
blockgroups['TRACTCE'] = blockgroups['TRACTCE'].apply('{0:0>6}'.format)
blockgroups['LATITUDE'] = blockgroups['LATITUDE'].astype(float)
blockgroups['LONGITUDE'] = blockgroups['LONGITUDE'].astype(float)

# filter to nyc
nyc_counties = ["081", "047", "061", "005", "085"]
blockgroups = blockgroups[blockgroups.COUNTYFP.isin(nyc_counties)]
```

@@

point_samples <- linestrings %>%
  # sample points along the line, one point per 500 feet.
  st_line_sample(density = 1/500)

@@

@@

```{r}

@@

>>>>

import numpy as np
import pandas as pd
import time
import multiprocessing as mp
```

@@

We use the `osmnx` library to get a highway network graph for New York City from 
OpenStreetMap. We get the walk network because most people in New York City will
not be driving to get to their park.

```{python graph}
graph = ox.graph_from_place('New York, New York, USA', network_type='drive')
graph_proj = ox.project_graph(graph)
```

@@

# pad the county and tract strings, and get the proper data types
blockgroups.STATEFP = blockgroups.STATEFP.apply(str)
blockgroups.BLKGRPCE = blockgroups.BLKGRPCE.apply(str)
blockgroups.COUNTYFP = blockgroups.COUNTYFP.apply('{0:0>3}'.format)
blockgroups.TRACTCE = blockgroups.TRACTCE.apply('{0:0>6}'.format)
blockgroups['GEOID'] = blockgroups[['STATEFP', 'COUNTYFP', 'TRACTCE', 'BLKGRPCE']].apply(lambda x: ''.join(x), axis = 1)

blockgroups['LATITUDE'] = blockgroups['LATITUDE'].astype(float)
blockgroups['LONGITUDE'] = blockgroups['LONGITUDE'].astype(float)

# filter to nyc
nyc_counties = ["081", "047", "061", "005", "085"]
blockgroups = blockgroups[blockgroups.COUNTYFP.isin(nyc_counties)]
blockgroups.set_index('GEOID', inplace=True)
```

@@

point_samples <- linestrings %>%
  # sample points along the line, one point per 1000 feet.
  st_line_sample(density = 1/1000)

@@

The results of this data munging are given below. We have some issues where parks 
made of multiple polygons end up with internal points. Trying to get rid of these
by using complex hulls created other problems on linear parks. So we'll just go
with this.

@@

Get the data into a form that we can pass it into python.

```{r park_points}
park_points <- open_space_points %>%
  st_transform(4326) %>%
  mutate(
    LONGITUDE = st_coordinates(.)[, 1],
    LATITUDE = st_coordinates(.)[,2]
  ) %>%
  st_set_geometry(NULL) %>%
  as_tibble()
```

## Distances

First, we bring the park points we sampled into a pandas data frame with
the park id as the index.
```{python park_points_py}
park_points = pd.DataFrame(r.park_points)
# TODO: run for all park points
park_ids = park_points.sample(5).id

park_points.set_index('id', inplace=True)
park_points = park_points.loc[park_ids]

@@

We already created the block group centroids data frame earlier. Let's get the
nearest network nodes for each of the block groups and each park point.

```{python park_bg_nodes}
# TODO: run for all block groups
blockgroups = blockgroups.sample(5)

def findnode(df):
  df['node'] = df.apply(lambda row: 
    ox.get_nearest_node(graph, (row.LATITUDE, row.LONGITUDE)), axis=1)
  return df

def parallelize_dataframe(df, func, n_cores=4):
  df_split = np.array_split(df, n_cores)
  pool = Pool(n_cores)
  df = pd.concat(pool.map(findnode, df_split))
  pool.close()
  pool.join()
  return df

a = time.process_time()
findnode(blockgroups)
findnode(park_points)
b = time.process_time()
b - a
```

We want to create a data structure that we can quickly populate with the 
shortest path between a block group and the nearest park point. This is 
probably a two-level dictionary.

```{python blockgroup_dict}
# Create a dictionary for the block groups
blockgroup_dict = {}
for i in blockgroups.index.unique():
  blockgroup_dict[i] = {} # create a dictionary for the parks
  for j in park_points.index.unique():
    blockgroup_dict[i][j] = float("inf")

Now we want to loop through each block group and all of the park points to figure
out what the shortest path is.

```{python nearest}
# loop through block groups
a = time.process_time()
for bg in blockgroups.itertuples():

  # loop through parks
  for park in park_ids:
    these_points = park_points.loc[[park]]

    # loop through points
    for point in these_points.itertuples():
      # get length between park and this edge point
      length = nx.shortest_path_length(graph_proj, source = bg.node, target=point.node,
        weight = 'length')

      # check if this is the shortest one we've found; if so, update.
      if length < blockgroup_dict[bg.Index][park]:
        blockgroup_dict[bg.Index][park] = length

b = time.process_time()
b - a 
```
